---
knit: (function(inputFile, encoding) { 
      out_dir <- '../../docs/distributions';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), out_dir, 'distributions.html')) })

output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Distributions

[Home](../index.html)

## Contents{#contents}



#### [Normal Distribution](#normalDistribution)
#### [Binomial Distribution](#binomialDistribution)
#### [Chi Squared Distribution](#chiSquaredDistribution)
#### [Student's t Distribution](#tDistribution)
#### [F Distribution](#fDistribution)

## Normal Distribution {#normalDistribution}

[contents](#contents)

### Probability Density

The Probability Density Function of the normal distribution is:

$$ f(x | \mu,\sigma^2) = \frac{1}{\sqrt{2\pi\sigma^2}}e^{-\frac{(x-\mu)^2}{2\sigma^2}}$$
Where:

 - $\mu$ is the [mean](https://en.wikipedia.org/wiki/Mean) or [expectation](https://en.wikipedia.org/wiki/Expected_value) of the distribution
  
 - $\sigma$ is the [standard deviation](https://en.wikipedia.org/wiki/Standard_deviation)
  
 - $\sigma^2$ is the [variance](https://en.wikipedia.org/wiki/Variance)

```{r normalDistribution}
x <- seq(-5, 5, by=0.1)
y <- dnorm(x)
plot(x,y, main="Standard Normal Distribution")
plot(x,y, type='l', main="Standard Normal Distribution Connected")
y <- dnorm(x, mean=2.5, sd=0.2)
plot(x,y, main="SND mean 2.5 , standard deviation 0.2")
lines(x,y, type="l")# lines adds to the existing plot without making a new one
```

### Cumulative Distribution

```{r normalDistributionCDF}
x = seq(-2, 2, by=0.01)
y = pnorm(x)
plot(x,y, main="Standard Normal Cumulative Distribution Function", type="l")
```

### Problems

1. Let X ~ N(5,3). Find P(X < 6)

```{r normalProblem1}
pnorm(6, mean=5, sd=sqrt(3))
```

2. Find P(4 < X < 6)

```{r normalProblem2}
pnorm(6, mean=5, sd=sqrt(3)) - pnorm(4, mean=5, sd=sqrt(3))
```

## Binomial Distribution {#binomialDistribution}
[contents](#contents)
```{r binomialDistribution}
x <- seq(0, 50, by=1)
y <- dbinom(x, size=50, prob=0.2)
plot(x,y, main="Binomial Distribution size 50, probability 0.2")
y <- dbinom(x, size=50, prob=0.6)
plot(x,y, main="Binomial Distribution size 50, probability 0.6")
y <- dbinom(x, size=100, prob=0.6)
plot(x,y, main="Binomial Distribution size 100, probability 0.6")
```

## Chi Squared Distribution {#chiSquaredDistribution}
[contents](#contents)

Let $X_1,X_2,...,X_n$ be independant and identically distributed random variables with an $\mathcal{N}(0,1)$ distribution (Standard Normal)

The Chi Square relates to the normal distribution in the following way:

$$ \sum_{i=1}^{n} X_i^2$$

has a $\chi^2$ distribution with n degrees of freedom and a variance of 2n


```{r chiSquaredDistribution}
x <- seq(0, 40, by=0.1)
y <- dchisq(x,1)
plot(x,y, main=" Chi Squared with 1 Degree of Freedom")
y <- dchisq(x,4)
plot(x,y, main=" Chi Squared with 4 Degree of Freedom")
y <- dchisq(x,10)
plot(x,y, main=" Chi Squared with 10 Degrees of Freedom")
y <- dchisq(x,20)
plot(x,y, main=" Chi Squared with 20 Degrees of Freedom")
```

## Student's t Distribution {#tDistribution}
[contents](#contents)

Note the tails are heavier then that of the normal distribution

```{r tDistribution}
x <- seq(-5, 5, by=0.1)
y <- dt(x, 1)
plot(x,y, main="t Distribution with 1 Degree of Freedom")
y <- dt(x, 10)
plot(x,y, main="t Distribution with 10 Degrees of Freedom")
y <- dt(x, 10000)
plot(x,y, main="t Distribution with 10000 Degrees of Freedom")
```

## F Distribution {#fDistribution}
[contents](#contents)

```{r fDistribution}
x <- seq(0, 10, by=0.05)
y <- df(x, 1, 1)
plot(x,y, main="F Distribution with 1 and 1 Degrees of Freedom")
y <- df(x, 100, 1)
plot(x,y, main="F Distribution with 100 and 1 Degrees of Freedom")
y <- df(x, 1, 100)
plot(x,y, main="F Distribution with 1 and 100 Degrees of Freedom")
y <- df(x, 100, 100)
plot(x,y, main="F Distribution with 100 and 100 Degrees of Freedom")
```