---
knit: (function(inputFile, encoding) { 
      out_dir <- '../../docs/examples';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), out_dir, 'misc.html')) })
                        
output: html_document
---

# Misc

[Home](../index.html) | [Examples](examples.html)

Adapted from University of Toronto, Mark Ebden, STA302

### Basic Plot With Linear Regression

Note: These are perhaps not the best x and y labels to compare.
```{r plot}
M <- c(102.5, 104.5, 100.4, 95.9, 87.0, 95.0, 88.6, 89.2,
78.9, 84.6, 81.7, 72.2, 65.1, 68.1, 67.3, 52.5)
T <- c(51.3, 49.9, 50.0,49.2, 48.5, 47.8, 47.3, 45.1,
46.3, 42.1, 44.2, 43.5, 42.3, 40.2, 31.8, 34.0)
plot(T,M,xlab="temperature",ylab="mortality index")
myFit <- lm(M~T) # M is the response variable and T the predictor
abline(myFit) # Add regression line to the plot

```

### Shade t Distribution

```{r shadeTDistribution}
c1 = qt(0.025,2) # Left bound of shaded region
c2 = qt(0.975,2)
x0 = 8 # Highest t-score to plot
myseq = seq(c1, c2, 0.01)
cx <- c(c1,myseq,c2) # vector of x-points to outline shaded region
cy <- c(0,dt(myseq,2),0)
curve(dt(x,2),xlim=c(-x0,x0),xlab='t',ylab='p(t)')
polygon(cx,cy,col='skyblue') # connect the dots

```

### Regression Line with Confidence Interval

Data: (x,y) ((2, 1),(4, 3),(6, 6))

The upper and lower limits of the Confidence Interval (CI) are given by the following formula,where:

 - t is the t distribution
 - Using 95% Confidence Interval
 
Variance of x:
 
$$ S_{xx} = \sum_{i=1}^n (x_i - \bar{x})^2  $$

Covariance of x and y:

$$ S_{xy} = \sum_{i=1}^n (x_i - \bar{x})(y_i - \bar{y})  $$

$$ b_1 = \frac{S_{xy}}{S_{xx}}$$

$$ b_0 = \frac{1}{n}\sum_{i=1}^n (y_i - b_1x_i) =\frac{1}{n}\sum_{i=1}^n (y_i) - b_1\frac{1}{n}\sum_{i=1}^n x_i = mean(\textbf{y}) - b_1mean(\textbf{x})$$

Residual Sum of Squares (RSS): 

$$ RSS = \sum_{i=1}^n \hat{e}_i^2 = \sum_{i=1}^n (y_i -b_0 - b_1x_i)^2$$

Mean Square Error:

$$ S^2 = MSE = \frac{RSS}{n-2}$$

CI: 

$$ \hat{y}^* = t(0.025, n-2) S \sqrt{\frac{1}{n} + \frac{(x^* - \bar{x})^2}{S_{xx}}}  $$

```{r regressionCI}
x<-c(2,4,6); y<-c(1,3,6); n <- length(x)
mx <- mean(x); my <- mean(y)
Sxx <- sum((x-mx)^2); Sxy <- sum((x-mx)*(y-my))
b1 <- Sxy/Sxx; b0 <- mean(y) - b1*mean(x)
yhat <- b0 + b1*x
RSS <- sum((y-yhat)^2)
S <- sqrt(RSS/(n-2))
xstar <- seq(min(x)-1,max(x)+1,.1) # Points at which to interpolate
ystarMean <- b0+b1*xstar # Interpolations
a <- qt(.975,n-2)*S*sqrt(1/n+(xstar-mx)^2/Sxx) # See slide 14
ystarLow <- ystarMean-a; ystarHigh <- ystarMean+a # Slide 14
plot(x,y,xlim=c(min(xstar),max(xstar)),
ylim=c(min(ystarLow),max(ystarHigh)))
lines(xstar,ystarMean,type="l",col="black")
lines(xstar,ystarLow,type="l",col="red")
lines(xstar,ystarHigh,type="l",col="red")
```