---
knit: (function(inputFile, encoding) { 
      out_dir <- '../../docs/examples';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), out_dir, 'twoSampletTest.html')) })
                        
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Two Sample t Test

[Home](../index.html) | [Examples](examples.html)

## Problem

Example Adapted from: https://onlinecourses.science.psu.edu/stat500/node/50

We are given two sets of samples. `new` and `old`. What we want to do is figure out if there is a difference in the mean of their distributions. More accurately, whether the mean of `new` is greater than the mean of `old`.

```{r twoSampletTest1}
new = c(42.1, 41.3, 42.4, 43.2, 41.8, 41.0, 41.8, 42.8, 42.3, 42.7) # Create a vector of samples
new # Display the vector to the console
old = c(42.7, 43.8, 42.5, 43.1, 44.0, 43.6, 43.3, 43.5, 41.7, 44.1)
old
n1 = length(new)
n2 = length(old)
```

## Sample Mean

New Sample Mean: $\bar{x}_1 = \frac{1}{n_1} \sum_{i=1}^{n_1} X_{i-new}$

Old Sample Mean: $\bar{x}_2 = \frac{1}{n_2} \sum_{i=1}^{n_2} X_{i-old}$

```{r twoSampletTest2}
newSampleMean = sum(new)/n1 # Using the equation for sample mean
newSampleMean
newSampleMean <- mean(new) # Easier to just use the R function
newSampleMean
oldSampleMean = mean(old)
oldSampleMean
```

$\bar{x}_1 = 42.14$ and $\bar{x}_2 = 43.23$

## Sample Variance and Standard Deviation

$\sigma_1^2 = \frac{1}{n_1 - 1} \sum_{i=1}^{n_1} (\bar{x}_1 - X_{i-new})$

$\sigma_2^2 = \frac{1}{n_2 - 1} \sum_{i=1}^{n_2} (\bar{x}_2 - X_{i-old})$

```{r twoSampletTest3}
newSampleVariance = sum( (newSampleMean - new) ^ 2 )/(n1 - 1) # Using Formula
newSampleVariance
newSampleVariance <- var(new) # Using R function
newSampleVariance
newStandardDeviation = sqrt(newSampleVariance)
newStandardDeviation
oldSampleVariance = var(old)
oldSampleVariance
oldStandardDeviation = sqrt(oldSampleVariance)
oldStandardDeviation
```

New Sample Variance: $\sigma_1^2 = 0.4671111$

New Standard Deviation: $\sigma_1 = 0.6834553$

Old Sample Variance: $\sigma_2^2 = 0.5623333$

Old Standard Deviation: $\sigma_2 = 0.7498889$

## Verify Normality

In order to be able to use the Two Sample t Test we need to verify that both samples follow a normal distribution. 

The following graphs show how closely the points line up on the actual normal distribution, they should fit as closely as possible.

The dotted lines represent the bounds for a 95% confidence interval. As we can see none of the points fall outside the 95% CI, meaning we can be confident that `old` and `new` come from a normal population.


```{r twoSampletTest4}
library(car) # Importing the car library which includes qqPlot
qqPlot(new)
qqPlot(old)
```

## Comparing Variance

In order to use the pooled t-test we must make sure the variance of the two samples is not too different. 

We compare the standard deviations, and want them to have a ratio between 0.5 and 2 (Elaborate more on why this is the case).

```{r twoSampletTest5}
newStandardDeviation/oldStandardDeviation
```

The standard deviation ratio is within 0.5 and 2, therefore we can perform the pooled t-test

## Pooled t-test

### Hypothesis

Let $\mu_1$ be the mean of `new` and $\mu_2$ be the mean of `old`

Then the hypotheses are:

$H_0 : \mu_1 - \mu_2 = 0,$

$H_a : \mu_1 - \mu_2 < 0$

### Significance Level

$\alpha = 0.05$

```{r setAlpha}
alpha = 0.05
```

### Compute the t-statistic

Estimate the common standard deviation using the pooled standard deviation:

$$s_p = \sqrt{\frac{(n_1 - 1)s_1^2 + (n_2 - 1)s_2^2}{n_1 + n_2 - 2}} $$

Test Statistic:

$$ t^* = \frac{\bar{x_1} - \bar{x_2}}{s_p\sqrt{\frac{1}{n_1} + \frac{1}{n_2}}}$$

with degrees of freedom $df = n_1 + n_2 - 2$

Plug in the variables:

$$s_p = \sqrt{\frac{(10 - 1)(0.6834553) + (10 - 1)(0.7498889)}{10 + 10 - 2}} $$

```{r twoSampletTest6}
sp = sqrt(((n1 - 1) * newSampleVariance + (n2 - 1) * oldSampleVariance)/(n1 + n2 - 2))
sp
```

$sp = 0.7174414$

$$ t^* = \frac{42.14 - 43.23}{(0.7174414)\sqrt{\frac{1}{10} + \frac{1}{10}}}$$

```{r twoSampletTest7}
tStatistic = (newSampleMean - oldSampleMean)/(sp * sqrt(1/n1 + 1/n2))
tStatistic
```

$t^* = -3.397231$

### Critical Value

Left Tailed Test (Explain why)

Critical Value: $-t_{\alpha} = -t_{0.05}$

Degrees of Freedom = $df = n1 + n2 - 2 = 10 + 10 - 2 = 18$

```{r twoSampletTest8}
df = n1 + n2 - 2
criticalValue = qt(alpha, df)
criticalValue
```

Rejection Region $t* < -1.734064

### Check Test Statistic vs Rejection Region

$t^* = -3.397231 < -1.734064$

```{r twoSampletTest9}
tStatistic < criticalValue
```

Therefore we reject $H_0$ at $\alpha = 0.05$

### Conclusion 

Therefore at the 5% significance level the mean of `new` is greater than the mean of `old`