---
title: "Distributions"
author: "NikitaVr"
date: "September 9, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[Home](index.html)

## Contents{#contents}

#### [Normal Distribution](#normalDistribution)
#### [Binomial Distribution](#binomialDistribution)
#### [Chi Squared Distribution](#chiSquaredDistribution)
#### [Student's t Distribution](#tDistribution)
#### [F Distribution](#fDistribution)

## Normal Distribution {#normalDistribution}

[contents](#contents)

### dnorm

```{r normalDistribution}
x <- seq(-5, 5, by=0.1)
y <- dnorm(x)
plot(x,y, main="Standard Normal Distribution")
plot(x,y, type='l', main="Standard Normal Distribution Connected")
y <- dnorm(x, mean=2.5, sd=0.2)
plot(x,y, main="SND mean 2.5 , standard deviation 0.2")
```

### Problems

1. Let X ~ N(5,3). Find P(X < 6)

```{r normalProblem1}
pnorm(6, mean=5, sd=sqrt(3))
```

2. Find P(4 < X < 6)

```{r normalProblem2}
pnorm(6, mean=5, sd=sqrt(3)) - pnorm(4, mean=5, sd=sqrt(3))
```

## Binomial Distribution {#binomialDistribution}
[contents](#contents)
```{r binomialDistribution}
x <- seq(0, 50, by=1)
y <- dbinom(x, size=50, prob=0.2)
plot(x,y, main="Binomial Distribution size 50, probability 0.2")
y <- dbinom(x, size=50, prob=0.6)
plot(x,y, main="Binomial Distribution size 50, probability 0.6")
y <- dbinom(x, size=100, prob=0.6)
plot(x,y, main="Binomial Distribution size 100, probability 0.6")
```

## Chi Squared Distribution {#chiSquaredDistribution}
[contents](#contents)
```{r chiSquaredDistribution}
x <- seq(0, 40, by=0.1)
y <- dchisq(x,1)
plot(x,y, main=" Chi Squared with 1 Degree of Freedom")
y <- dchisq(x,4)
plot(x,y, main=" Chi Squared with 4 Degree of Freedom")
y <- dchisq(x,10)
plot(x,y, main=" Chi Squared with 10 Degrees of Freedom")
y <- dchisq(x,20)
plot(x,y, main=" Chi Squared with 20 Degrees of Freedom")
```

## Student's t Distribution {#tDistribution}
[contents](#contents)

Note the tails are heavier then that of the normal distribution

```{r tDistribution}
x <- seq(-5, 5, by=0.1)
y <- dt(x, 1)
plot(x,y, main="t Distribution with 1 Degree of Freedom")
y <- dt(x, 10)
plot(x,y, main="t Distribution with 10 Degrees of Freedom")
y <- dt(x, 10000)
plot(x,y, main="t Distribution with 10000 Degrees of Freedom")
```

## F Distribution {#fDistribution}
[contents](#contents)

```{r fDistribution}
x <- seq(0, 10, by=0.05)
y <- df(x, 1, 1)
plot(x,y, main="F Distribution with 1 and 1 Degrees of Freedom")
y <- df(x, 100, 1)
plot(x,y, main="F Distribution with 100 and 1 Degrees of Freedom")
y <- df(x, 1, 100)
plot(x,y, main="F Distribution with 1 and 100 Degrees of Freedom")
y <- df(x, 100, 100)
plot(x,y, main="F Distribution with 100 and 100 Degrees of Freedom")
```